
(This code was developed based on Aleah's 02_emailed studies code.) 

The aim for this code is to 1) generate a list of arrays (studies) that contain matrices of each study (studies have different number of populations and matrix dimensions).  and 2) to create a matrix that is latitude and longitude in columns, with rows being each unique study.


## 0. load packages
```{r}
library(dplyr)
library(readxl)
library(abind)
library(rio)

```


## 1. create study name vectors
```{r}
vector_studies <-list.files( path="F:/Sensitivity meta analysis/sen-across-space/Raw data") #list of studies, not including path, including ".xlsx"
vector_studies_noxl<- gsub(".xlsx", "", vector_studies) # *********This is the character vector of study names for Allison*************
vector_studies_WITHPATH <-list.files("F:/Sensitivity meta analysis/sen-across-space/Raw data", full.names = TRUE) #this vector is useful for reading in files with for() loop

```

## 2. load studies into a list
```{r}
study_list <- lapply(vector_studies_WITHPATH, import_list, col_names = FALSE)  ## use import_list to load all tabs at the same time. Loop over all studies and add them (each one being a list) to a large list
names(study_list) <- vector_studies

```

## 3. create an array for each study
```{r}
study1 <- study_list[[1]]

num_pop <- length(study1)-2
mtx_dim <- dim(study1[[3]])[1]## number of population: number of tabs besides README and info
array1 <- array(NA, dim = c(mtx_dim, mtx_dim, num_pop))

for (i in 1:num_pop) {
  j <- i+2
  array1[, , i] <- as.matrix(study1[[j]])
} 





```

