
(This code was developed based on Aleah's 02_emailed studies code.) 

The aim for this code is to 1) generate a list of arrays (studies) that contain matrices of each study (studies have different number of populations and matrix dimensions).  and 2) to create a matrix that is latitude and longitude in columns, with rows being each unique study.


## 0. load packages
```{r}
library(dplyr)
library(readxl)
library(abind)
library(rio)
library(openxlsx)

```


## 1. create study name vectors
```{r}
vector_studies <-list.files( path="F:/Sensitivity meta analysis/sen-across-space/Raw data") #list of studies, not including path, including ".xlsx"
vector_studies_noxl<- gsub(".xlsx", "", vector_studies) # *********This is the character vector of study names for Allison*************
vector_studies_WITHPATH <-list.files("F:/Sensitivity meta analysis/sen-across-space/Raw data", full.names = TRUE) #this vector is useful for reading in files with for() loop

```

## 2. load studies into a list
```{r}
study_list <- lapply(vector_studies_WITHPATH, import_list, col_names = FALSE)  ## use import_list to load all tabs at the same time. Loop over all studies and add them (each one being a list) to a large list
## col_names=False to keep the first row of the matrix
names(study_list) <- vector_studies

```

#### a. average across years

Munzgergova et al., 2005 matrices are not in correct format, have to average them across years. m = (m1+m2+m3)/3

Start with the first paper
```{r}
names(study_list)

Munz_1 <- study_list[[10]]

Munz_1_A <- as.matrix(Munz_1[[3]])
m1 <- as.matrix(Munz_1_A[3:5, 4:6])
m2 <- as.matrix(Munz_1_A[8:10, 4:6])
m3 <- as.matrix(Munz_1_A[13:15, 4:6])
class(m1) <- class(m2)<- class(m3) <- "numeric"   ## need to convert these character matrices into numeric to perform element wise calculation
Encovany <- (m1+m2+m3)/3

Munz_1_B <- as.matrix(Munz_1[[4]])
m1 <- as.matrix(Munz_1_B[3:5, 4:6])
m2 <- as.matrix(Munz_1_B[8:10, 4:6])
m3 <- as.matrix(Munz_1_B[13:15, 4:6])
class(m1) <- class(m2)<- class(m3) <- "numeric"
Malešov <- (m1+m2+m3)/3

Munz_1_C <- as.matrix(Munz_1[[5]])
m1 <- as.matrix(Munz_1_C[3:5, 4:6])
m2 <- as.matrix(Munz_1_C[8:10, 4:6])
m3 <- as.matrix(Munz_1_C[13:15, 4:6])
class(m1) <- class(m2)<- class(m3) <- "numeric"
Malíč <- (m1+m2+m3)/3

### add the averaged matrices back to the Munz_1 list
Munz_1[["Encovany"]] <- as.data.frame(Encovany)
Munz_1[["Malešov"]] <- as.data.frame(Malešov)
Munz_1[["Malíč"]] <- as.data.frame(Malíč)

### delete the unprocessed list elements
Munz_1 <- Munz_1[!names(Munz_1) %in% c("A", "B", "C")]

```

Then the second paper (same issue)
```{r}

Munz_2 <- study_list[[11]]

Munz_2_A <- as.matrix(Munz_2[[3]])
m1 <- as.matrix(Munz_2_A[3:5, 4:6])
m2 <- as.matrix(Munz_2_A[8:10, 4:6])
m3 <- as.matrix(Munz_2_A[13:15, 4:6])
class(m1) <- class(m2)<- class(m3) <- "numeric"   ## need to convert these character matrices into numeric to perform element wise calculation
Hovr <- (m1+m2+m3)/3

Munz_2_B <- as.matrix(Munz_2[[4]])
m1 <- as.matrix(Munz_2_B[3:5, 4:6])
m2 <- as.matrix(Munz_2_B[8:10, 4:6])
m3 <- as.matrix(Munz_2_B[13:15, 4:6])
class(m1) <- class(m2)<- class(m3) <- "numeric"
Loreta <- (m1+m2+m3)/3

Munz_2_C <- as.matrix(Munz_2[[5]])
m1 <- as.matrix(Munz_2_C[3:5, 4:6])
m2 <- as.matrix(Munz_2_C[8:10, 4:6])
m3 <- as.matrix(Munz_2_C[13:15, 4:6])
class(m1) <- class(m2)<- class(m3) <- "numeric"
Malíč <- (m1+m2+m3)/3

### add the averaged matrices back to the Munz_2 list
Munz_2[["Hovr"]] <- as.data.frame(Hovr)
Munz_2[["Loreta"]] <- as.data.frame(Loreta)
Munz_2[["Malíč"]] <- as.data.frame(Malíč)

### delete the unprocessed list elements
Munz_2 <- Munz_2[!names(Munz_2) %in% c("A", "B", "C")]
```

Add the updated list back to study_list
```{r}

study_list[["Munzgergova 2005 Cirsium acaule.xlsx"]] <- Munz_1  
study_list[["Munzgergova 2005 Cirsium pannonicum.xlsx"]] <- Munz_2
```


#### b. check for consistent matrix dimensions

Check whether there're studies with variable matrix dimensions. For the following codes to run, each population tab should have the same dimension. 

```{r}


study_list_same_dim <- list()
study_list_var_dim <- list()

for (i in 1:length(study_list)) {
  
  study_i <- study_list[[i]]
  dims_i <- sapply(study_i[3:length(study_i)], dim)
  
  if (length(unique(dims_i[1, ])) == 1) {
    
    study_list_same_dim[[length(study_list_same_dim) + 1]] <- study_i
    names(study_list_same_dim)[length(study_list_same_dim)] <- names(study_list)[i]
    
  } else {
       study_list_var_dim[[length(study_list_var_dim) + 1]] <- study_i
       names(study_list_var_dim)[length(study_list_var_dim)] <- names(study_list)[i]
  }
}

```

Doak & Morris 2010 Polygonum viviparum.xlsx have populations of variable matrix dimensions. Need to consult with Allison on how to deal with it.


## 3. create an array for each study (only if it contains mtx with a same dimension)
```{r}
mtx_array_study_list <-  vector("list", length = length(study_list_same_dim))

for ( k in 1:length(study_list_same_dim)) {
  study_k <- study_list_same_dim[[k]]
  num_pop_k <- length(study_k)-2
  mtx_dim_k <- dim(study_k[[3]])[1]## number of population: number of tabs besides README and info
  array_k <- array(NA, dim = c(mtx_dim_k, mtx_dim_k, num_pop_k))
  
  for (i in 1:num_pop_k) {
  j <- i+2
  array_k[, , i] <- as.matrix(study_k[[j]])
  } 
  
  mtx_array_study_list[[k]] <- array_k
}

names(mtx_array_study_list) <- names(study_list_same_dim)

matrices_emailed_1 <- mtx_array_study_list

```

## 4. create a location matrix

extract all the info tabs

```{r}
info_list <- list()

for (k in 1:length(study_list)) {
  study_k <- study_list[[k]]
  
  info_k <- as.data.frame(study_list[[k]][names(study_list[[k]]) %in% c("info","Info", "INFO")])
  
  info_list[[k]] <- info_k
}

names(info_list) <- vector_studies_noxl

```

```{r}
## run this before the loop, this function check for the first row of the location matrix
start_row <- function(x) {  
    any(grepl("population", x, ignore.case = TRUE)) &
    any(grepl("\\blat",        x, ignore.case = TRUE)) & ## lat has to be in the beginning of a phrase (e.g. not in the word "population")
    any(grepl("\\blong",       x, ignore.case = TRUE))}


location_list <-list() 
studies_withno_coord <- list()


for (i in 1: length(info_list)) {
  study_i <- info_list[[i]] 
  
  if (any(grepl("\\b(lat|long)", as.matrix(study_i), ignore.case = TRUE))) {header <- which(apply(study_i, 1, start_row))[1] ## extract the row num of the first row that satisfies the conditions
  
  study_i_2 <- study_i[header,]   ## extract col position for the three columns we need
  pop_col <- which(grepl("population", study_i_2[1,], ignore.case = TRUE))[1]
  lat_col <- which(grepl("\\blat", study_i_2[1,], ignore.case = TRUE))[1]
  long_col <- which(grepl("\\blong", study_i_2[1,], ignore.case = TRUE))[1]
  loc_mtx_untrimmed <- study_i[
    (header+1):nrow(study_i),
    c(pop_col, lat_col, long_col),
    drop = FALSE]
  names(loc_mtx_untrimmed) <- c("Population", "Lattitude", "Longitude")
  
  ## trim white space and add study name to pop names
  if (any(is.na(loc_mtx_untrimmed$Population))) {
  first_NA <- which(is.na(loc_mtx_untrimmed$Population))[1]
  loc_mtx_i <- loc_mtx_untrimmed[-c(first_NA:nrow(loc_mtx_untrimmed)) ,]} else 
  {loc_mtx_i<-loc_mtx_untrimmed}
  
  study_name <- vector_studies_noxl[i]
  # loc_mtx_i$Population <- paste(loc_mtx_i$Population, study_name, sep = ", ")
  
  location_list [[study_name]] <- as.data.frame (loc_mtx_i)}  else {  ## export bad studies with no lat/long in Info
    study_name_bad <- vector_studies_noxl[i]
    studies_withno_coord[[study_name_bad]] <- study_i
    
}
  }



    
```

bind location_list into a single matrix. Clean up.

```{r}
locations_emailed <- bind_rows(location_list, .id = "study name")
```

Delete E, N. Keep only numbers.

```{r}
locations_emailed$Lattitude <- gsub('[N|S]', '', locations_emailed$Lattitude)
locations_emailed$Longitude <- gsub('[E|W]', '', locations_emailed$Longitude)

```

Fix reversed Lat/long

```{r}
wrong_lat <- locations_emailed$Lattitude[which(locations_emailed$`study name` %in% c("Herzog et al. in review Sedum lanceolatum"))]
wrong_long <- locations_emailed$Longitude[which(locations_emailed$`study name` %in% c("Herzog et al. in review Sedum lanceolatum"))]

## swap the lat and long
locations_emailed$Longitude[which(locations_emailed$`study name` %in% c("Herzog et al. in review Sedum lanceolatum"))] <- wrong_lat
locations_emailed$Lattitude[which(locations_emailed$`study name` %in% c("Herzog et al. in review Sedum lanceolatum"))] <- wrong_long

```



## 5. export
At the end this code should export 1. matrices_emailed, 2. location_emailed, and 3. studyID_emailed

```{r}
studyID_emailed <- vector_studies_noxl
write.csv(studyID_emailed, "processed data/studyID_emailed.csv", row.names = FALSE)

save(matrices_emailed_1, file = "processed data/matrices_emailed_1.RData")

write.csv(locations_emailed, file = "processed data/locations_emailed.csv")

```

Export all three as one file for TODO list 1.

```{r}
save(matrices_emailed_1, locations_emailed, studyID_emailed, file = "processed data/TODO_list_01_output.RData" )
```




