
(This code was developed based on Aleah's 02_emailed studies code.) 

The aim for this code is to 1) generate a list of arrays (studies) that contain matrices of each study (studies have different number of populations and matrix dimensions).  and 2) to create a matrix that is latitude and longitude in columns, with rows being each unique study.


## 0. load packages
```{r}
library(dplyr)
library(readxl)
library(abind)
library(rio)

```


## 1. create study name vectors
```{r}
vector_studies <-list.files( path="F:/Sensitivity meta analysis/sen-across-space/Raw data") #list of studies, not including path, including ".xlsx"
vector_studies_noxl<- gsub(".xlsx", "", vector_studies) # *********This is the character vector of study names for Allison*************
vector_studies_WITHPATH <-list.files("F:/Sensitivity meta analysis/sen-across-space/Raw data", full.names = TRUE) #this vector is useful for reading in files with for() loop

```

## 2. load studies into a list
```{r}
study_list <- lapply(vector_studies_WITHPATH, import_list, col_names = FALSE)  ## use import_list to load all tabs at the same time. Loop over all studies and add them (each one being a list) to a large list
## col_names=False to keep the first row of the matrix
names(study_list) <- vector_studies

```

#### a. average across years

Munzgergova et al., 2005 matrices are not in correct format, have to average them across years. m = (m1+m2+m3)/3


```{r}
names(study_list)

Munz_1 <- study_list[[10]]

Munz_1_A <- as.matrix(Munz_1[[3]])
m1 <- as.matrix(Munz_1_A[3:5, 4:6])
m2 <- as.matrix(Munz_1_A[8:10, 4:6])
m3 <- as.matrix(Munz_1_A[13:15, 4:6])
class(m1) <- class(m2)<- class(m3) <- "numeric"   ## need to convert these character matrices into numeric to perform element wise calculation
Encovany <- (m1+m2+m3)/3

Munz_1_B <- as.matrix(Munz_1[[4]])
m1 <- as.matrix(Munz_1_B[3:5, 4:6])
m2 <- as.matrix(Munz_1_B[8:10, 4:6])
m3 <- as.matrix(Munz_1_B[13:15, 4:6])
class(m1) <- class(m2)<- class(m3) <- "numeric"
Malešov <- (m1+m2+m3)/3

Munz_1_C <- as.matrix(Munz_1[[5]])
m1 <- as.matrix(Munz_1_C[3:5, 4:6])
m2 <- as.matrix(Munz_1_C[8:10, 4:6])
m3 <- as.matrix(Munz_1_C[13:15, 4:6])
class(m1) <- class(m2)<- class(m3) <- "numeric"
Malíč <- (m1+m2+m3)/3


```


#### b. check for consistent matrix dimensions

```{r}
## check whether there're studies with variable matrix dimensions

study_list_same_dim <- list()
study_list_var_dim <- list()

for (i in 1:length(study_list)) {
  
  study_i <- study_list[[i]]
  dims_i <- sapply(study_i[3:length(study_i)], dim)
  
  if (length(unique(dims_i[1, ])) == 1) {
    
    study_list_same_dim[[length(study_list_same_dim) + 1]] <- study_i
    names(study_list_same_dim)[length(study_list_same_dim)] <- names(study_list)[i]
    
  } else {
       study_list_var_dim[[length(study_list_var_dim) + 1]] <- study_i
       names(study_list_var_dim)[length(study_list_var_dim)] <- names(study_list)[i]
  }
}

## looks like there's one study with variable matrices
```





## 3. create an array for each study (only if it contains mtx with a same dimension)
```{r}
study_list_mtx_array <-  vector("list", length = length(study_list_same_dim))

for ( k in 1:length(study_list_same_dim)) {
  study_k <- study_list_same_dim[[k]]
  num_pop_k <- length(study_k)-2
  mtx_dim_k <- dim(study_k[[3]])[1]## number of population: number of tabs besides README and info
  array_k <- array(NA, dim = c(mtx_dim_k, mtx_dim_k, num_pop_k))
  
  for (i in 1:num_pop_k) {
  j <- i+2
  array_k[, , i] <- as.matrix(study_k[[j]])
  } 
  
  study_list_mtx_array[[k]] <- array_k
}









```

